\subsection{NFC (Near Field Communication)}
\label{subsection:nfc}

\gls{NFC} is a short-range high frequency wireless communication technology that allows data transfer between devices over small distances, first introduced by three manufacturers, Sony, Nokia and Philips. Communications maintain interoperability between other different communication methods, such as Bluetooth.

\gls{NFC} can be used to connect mobile applications with the physical world, \textit{e.g.} home appliances, connect devices through physical proximity, forming a peer-to-peer network, and card emulation, creating a connection to a common infrastructure and allowing some actions on the infrastructure, such as making payments.

Evolved from \gls{RFID} technology, an \gls{NFC} chip operates as one part of a wireless link. Once it is activated by another chip, small amounts of data can be exchanged between the pair if within a few centimeters from each other. One of the advantages, compared to other wireless communication technologies, is that \gls{NFC} does not requires a setup to pair two devices, thus reducing the time and packets exchanged in the transaction, allowing for times up to 1ms. Also, \gls{NFC} chips run on low amounts of power, making this technology much more power-efficient than other technologies.

The short range of the \gls{NFC} technology is a disadvantage due to its spacial limitation, but in terms of security, this spacial requirement provides a higher degree of security than Bluetooth. For instance, making \gls{NFC} relatively secure to use in crowded areas, where other wireless technologies could be impossible to use to transfer sensitive data, such as credit card data.

\gls{NFC} operates at 13.56MHz using Amplitude Shift Keying as the modulation scheme and \gls{TDD} for simultaneous receive and transfer of data, achieving data rates up to 424kbps, and although these rates are not impressive, for the amount of data that is sought to exchange using this technology and the lack of necessity for communication setup, \gls{NFC} provides relatively fast transfer times.






